### 2015-01-28 fitting.Rmd
Port from tree to tree2

```{r, echo=FALSE, results="hide"}
sourceDir <- function(path, ...) {
     for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
        source(file.path(path, nm), ...)
     }
}
sourceDir("R")
```

@@

```{r, echo=TRUE, results="hide"}
library(maker)
library(dplyr)

m <- maker::maker()
m$make("BCI_species_data")
data <- m$get("BCI_species_data")
head(data)
```

@@

```{r, echo=TRUE, results="hide"}
library(tree)

# given a dbh x, traits and parameters, grows plant to right size and estimates growth rate
run_tree <- function(x, E=1, traits=list(), pars=list(), strategy = default_strategy()) {

  distance_from_target_fn <- function(plant) {
    plant$vars_size[["diameter"]] - x
  }

  strategy <- strategy$copy()
  strategy$set_parameters(traits)
  strategy$set_parameters(pars)
  plant1 <- new(Plant, strategy)
  plant1$height <- 0.1
  plant <- try(grow.plant.to.size(plant1, fixed.environment(E), distance.from.target = distance_from_target_fn), silent = FALSE)
  if (inherits(plant, "try-error"))
      return(0)
  run_plant(plant, E)$dbasal_diam_dt
}
```
We can test it works

@@

```{r, echo=TRUE, results="show"}
library(minpack.lm)
## values over which to simulate data
x <- seq(0,5,length=100)
## model based on a list of parameters
getPred <- function(parS, xx) parS$a * exp(xx * parS$b) + parS$c
## parameter values used to simulate data
pp <- list(a=9,b=-1, c=6)
## simulated data, with noise
simDNoisy <- getPred(pp,x) + rnorm(length(x),sd=.1)
## plot data
plot(x,simDNoisy, main="data")
## residual function
residFun <- function(p, observed, xx) observed - getPred(p,xx)
## starting values for parameters
parStart <- list(a=3,b=-.001, c=1)
## perform fit
nls.out <- nls.lm(par=parStart, fn = residFun, observed = simDNoisy,
xx = x, control = nls.lm.control(nprint=1))
## plot model evaluated at final parameter estimates
lines(x,getPred(as.list(coef(nls.out)), x), col=2, lwd=2)
## summary information on parameter estimates
summary(nls.out)
```

@@

```{r, echo=TRUE, results="hide"}
# generative mdoel
# takes pars and traits
getPred <- function(pars, xx) {
	npoints <- nrow(xx)
	zz <- rep(NA_real_, npoints)
	for(i in seq_len(npoints)) {
		zz[i] <- run_tree(xx$dbh[i], E=1, traits=list(lma=xx$lma[i], rho=xx$rho[i], hmat=100), pars=pars)
	}
	zz
}
```

@@

```{r, echo=TRUE, results="hide"}
residFun <- function(p, observed, xx) observed - getPred(p,xx)
```

The set parameters to fit an some lower bounds (optionl) for these:

@@

```{r, echo=TRUE, results="show"}
x_example <- data[c(1,10), c("dbh", "lma", "rho")]
obs_example <- data[c(1,10), c("dbasal_diam_dt")]
getPred(par_init, x_example)
residFun(par_init, obs_example, x_example)
```

@@

nls.out <- nls.lm(par=par_init, fn = residFun, observed = obs, xx = x,
		lower=lower,
		control = nls.lm.control(nprint=1, ftol=1e-4))

out.init <- getPred(par_init, x)
out.fit <- getPred(as.list(coef(nls.out)), x)
```

@@

	out.init <- getPred(par_init, xs)
	out.fit <- getPred(as.list(coef(nls.out)), xs)

>>>>

```{r, echo=FALSE, results="hide"}
source_dir <- function(path, ...) {
  for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
    source(file.path(path, nm), ...)
  }
}
source_dir("R")
```

@@

```{r, echo=TRUE, results="hide"}
library(dplyr)

m <- maker::maker()
data <- m$make("BCI_species_data")
head(data)
```

@@

```{r, echo=TRUE, results="hide"}
library(tree2)

# given a dbh x, traits and parameters, grows plant to right size and estimates growth rate
run_tree <- function(x, E=1, traits=list(), pars=list(),
                     strategy=default_strategy()) {
  strategy[names(traits)] <- traits
  strategy[names(pars)] <- pars
  pp <- grow_plant_to_size(Plant(strategy), x, "diameter",
                           fixed_environment(E))
  sapply(pp$plant, function(p) p$vars_growth[["dbasal_diam_dt"]])
}
```
We can test it works

@@

```{r, echo=TRUE, results="show"}
library(minpack.lm)
## values over which to simulate data
x <- seq(0, 5, length.out=100)
## model based on a list of parameters
get_pred <- function(pars, xx) pars$a * exp(xx * pars$b) + pars$c
## parameter values used to simulate data
pp <- list(a=9, b=-1, c=6)
## simulated data, with noise
sim_d_noisy <- get_pred(pp,x) + rnorm(length(x), sd=.1)
## plot data
plot(x, sim_d_noisy, main="data")
## residual function
resid_fun <- function(p, observed, xx) observed - get_pred(p,xx)
## starting values for parameters
par_start <- list(a=3, b=-.001, c=1)
## perform fit
nls_out <- nls.lm(par=par_start, fn = resid_fun, observed = sim_d_noisy,
xx = x, control = nls.lm.control(nprint=1))
## plot model evaluated at final parameter estimates
lines(x,get_pred(as.list(coef(nls_out)), x), col=2, lwd=2)
## summary information on parameter estimates
summary(nls_out)
```

@@

```{r, echo=TRUE, results="hide"}
# generative mdoel
# takes pars and traits
get_pred <- function(pars, xx) {
  npoints <- nrow(xx)
  zz <- rep(NA_real_, npoints)
  for(i in seq_len(npoints)) {
    zz[i] <- run_tree(xx$dbh[i], E=1,
                      traits=list(lma=xx$lma[i], rho=xx$rho[i], hmat=100),
                      pars=pars)
  }
  zz
}
```

@@

```{r, echo=TRUE, results="hide"}
resid_fun <- function(p, observed, xx) observed - get_pred(p,xx)
```

The set parameters to fit an some lower bounds (optional) for these:

@@

```{r, echo=TRUE, results="show"}
x_example <- data[c(1,10), c("dbh", "lma", "rho")]
obs_example <- data[c(1,10), c("dbasal_diam_dt")]
get_pred(par_init, x_example)
resid_fun(par_init, obs_example, x_example)
```

@@

nls_out <- nls.lm(par=par_init, fn = resid_fun, observed = obs, xx = x,
                  lower=lower,
                  control = nls.lm.control(nprint=1, ftol=1e-4))

out.init <- get_pred(par_init, x)
out.fit <- get_pred(as.list(coef(nls_out)), x)
```

@@

  out.init <- get_pred(par_init, xs)
  out.fit <- get_pred(as.list(coef(nls_out)), xs)
